## Load dataset from YALE0065_OUTPUT_cleaned.xlsx file

```{r}

library(car)

# Load the data (update the path as needed)
library(readxl)
data <- read_excel("YALE0065_OUTPUT_cleaned.xlsx", sheet = "Combined sheet")
```

### *generation* and *educ* into factors and relabeling

```{r}
# Convert generation and educ to factors
data$generation <- as.factor(data$generation)
data$educ <- as.factor(data$educ)

# Relabel `generation` according to the schema
data$generation <- dplyr::recode(data$generation,
                          "z" = "Gen Z (1997–2012)",
                          "y" = "Millennials (1981–1996)",
                          "x" = "Gen X (1965–1980)",
                          "b" = "Boomers (1946–1964)",
                          "s" = "Silent generation (1928–1945)")

# Relabel `educ` according to the schema
data$educ <- dplyr::recode(data$educ,
                           "1" = "No HS",
                           "2" = "High school graduate",
                           "3" = "Some college",
                           "4" = "2-year",
                           "5" = "4-year",
                           "6" = "Post-grad")
```

### TWO-WAY ANOVA

```{r}
# Run the ANOVA
model <- aov(Tech_Average ~ generation * educ, data = data)

# View the summary of the results
summary(model)
```

### Compute the Effect Size 

```{r}
# SS values from the ANOVA summary
SS_generation <- 2.23
SS_educ <- 10.68
SS_interaction <- 1.01
SS_residual <- 149.22

# Calculate partial η² for generation, education, and interaction
partial_eta2_generation <- SS_generation / (SS_generation + SS_residual)
partial_eta2_educ <- SS_educ / (SS_educ + SS_residual)
partial_eta2_interaction <- SS_interaction / (SS_interaction + SS_residual)

# Print results
cat("Partial η² for generation:", partial_eta2_generation, "\n")
cat("Partial η² for education:", partial_eta2_educ, "\n")
cat("Partial η² for interaction:", partial_eta2_interaction, "\n")

```

###  

### Means and SD for Tech_Average by generation

```{r}
#Means cat("Means for Tech_Average by generation:\n") 
Tapply(Tech_Average ~ generation, mean, na.action=na.omit,         data=data) 
```

```{r}
# means  cat("\n\nStandard Deviations for EMOTION_Coping by Gender:\n") 
Tapply(Tech_Average ~ generation, sd, na.action=na.omit,            data=data) # stand dev
```

### Means and SD for Tech_Average by educ

```{r}
#Means cat("Means for Tech_Average by educ:\n") 
Tapply(Tech_Average ~ educ, mean, na.action=na.omit,         data=data) 
```

```{r}
# means  cat("\n\nStandard Deviations for Tech_Average by educ:\n") 
Tapply(Tech_Average ~ educ, sd, na.action=na.omit,            data=data) # stand dev
```

After finding significant results from **ANOVA** tests, we use a **post-hoc test** (**Tukey's HSD**) *to compare specific groups against each other and determine where the differences lie.*

```{r}
# Compare means of the groups, using Tukey-Kramer Post-hoc.
TukeyHSD(model)
```

### Check assumptions

```{r}
# Homogeneity test
leveneTest(Tech_Average ~ generation*educ,data= data, center="mean")
```

```{r}
# Perform Shapiro-Wilk test
shapiro.test(data$Tech_Average)
```

### Visualize results

```{r}
# Compute summary statistics for plotting
summary_data <- data %>%
  group_by(generation, educ) %>%
  summarise(mean = mean(Tech_Average, na.rm = TRUE),
            se = sd(Tech_Average, na.rm = TRUE) / sqrt(n()))

# Bar plot with error bars
ggplot(summary_data, aes(x = generation, y = mean, fill = educ)) +
  geom_bar(stat = "identity", position = "dodge") +
  geom_errorbar(aes(ymin = mean - se, ymax = mean + se), 
                position = position_dodge(0.9), width = 0.2) +
  labs(title = "Mean Tech_Average by Generation and Education",
       x = "Generation",
       y = "Mean Tech_Average",
       fill = "Education") +
  theme_minimal()


```

### Calculate and Plot the Estimated Marginal Means (EMM)

```{r}
# Run the two-way ANOVA
model <- aov(Tech_Average ~ generation * educ, data = data)

# Load emmeans for visualization
library(emmeans)

# Get estimated marginal means for the combination of generation and educ
emm <- emmeans(model, ~ generation * educ)

# Plot the estimated marginal means
plot(emm, comparisons = TRUE, type = "response", 
     main = "Estimated Marginal Means of Tech_Average by Generation and Education")


```
